---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
op <- options(warn = -1)

pl <- as.data.frame(rbind(
c(158.93835,-54.49871), 
c(77 + 58/60 + 3/3600,-(68 + 34/60 + 36/3600)),
c(62 + 52/60 + 27/3600, -(67 + 36/60 + 12/3600)),
c(110 + 31/60 + 36/3600,-(66 + 16/60 + 57/3600)),
c(73 + 30/60 + 30/3600,-(53 + 0 + 0/3600))))
names(pl) <- c("x", "y")
pl$name <- c("Macquarie Island", "Davis Station", "Casey Station", "Mawson Station", "Heard Island")
```
# seaice.map

<!-- badges: start -->
<!-- [![R-CMD-check](https://github.com/mdsumner/seaice.map/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mdsumner/seaice.map/actions/workflows/R-CMD-check.yaml)-->
<!-- badges: end -->

The goal of seaice.map is to 

- download very latest available sea ice concentration data
- generate a global map of the data with a customizable focus (i.e. global but centred on Tasmania/East Antarctica)
- also show some contextual spatio-temporal data (i.e. recent movements of a research vessel)
- explicate how to do this all with free tools. 

First, a modified map of the subsequent one to put the ship in the centre.  (we'll fix this up)

```{r pivot-map, fig.dim = c(6, 14.4), results = "show", echo=FALSE}
n <- 30 * 24 * 60

dat <- arrow::read_parquet("https://github.com/mdsumner/nuyina.underway/raw/main/data-raw/nuyina_underway.parquet")


print(range( dat$date_time_utc))
dat <- tail(dat, n)

track <- cbind(dat$longitude, dat$latitude)

library(terra)
r <- vapour::gdal_raster_data("data-raw/seaice.png", bands = 1:3, target_crs = sprintf("+proj=tmerc +lon_0=%f", track[nrow(track), 1]), 
                              target_dim = c(1024, 0))

pcrs <- attr(r, "projection")



ximage::ximage(r, asp = 1, axes = FALSE)

points(terra::project(do.call(cbind, maps::map(plot = F)[1:2]), to = pcrs, from = "OGC:CRS84"), pch = ".", col = "#777777")
title(readLines("data-raw/latestdate.txt"), line = -2, col.main = "white")



ptrack <- terra::project(as.matrix(track), to = pcrs, from = "OGC:CRS84")
lines(ptrack, col = "hotpink")
pt <- tail(ptrack[!is.na(ptrack[,1]) & !is.na(ptrack[,2]), ], 1L)
points(pt, pch = "+", col = "hotpink")
## key locations just defined in the source of this document

pl[c("X", "Y")] <- terra::project(cbind(pl$x, pl$y), to = pcrs, from = "OGC:CRS84")
points(pl$X, pl$Y, pch = 19, col = "hotpink", cex = 0.5)
bx <- c(range(track[,1], na.rm = TRUE), range(track[,2], na.rm = TRUE))
rect(bx[1], bx[3], bx[2], bx[4])

claims <- terra::project(terra::vect("data-raw/claims/claim_boundaries_ps.shp"), pcrs)
plot(claims, add = TRUE)

```


Now the map we carefully designed (but keep changing to figure out how we'd like to think about East Antarctica). 

```{r example, fig.dim = c(6, 14.4), results="show"}
library(terra)
r <- vapour::gdal_raster_data("data-raw/seaice.png", bands = 1:3)
pcrs <- attr(r, "projection")
ximage::ximage(r, asp = 1, axes = FALSE)
points(terra::project(do.call(cbind, maps::map(plot = F)[1:2]), to = pcrs, from = "OGC:CRS84"), pch = ".", col = "#777777")
title(readLines("data-raw/latestdate.txt"), line = -2, col.main = "white")

ptrack <- terra::project(as.matrix(track), to = pcrs, from = "OGC:CRS84")

lines(ptrack, col = "hotpink")
pt <- tail(ptrack[!is.na(track[,1]) & !is.na(ptrack[,2]), ], 1L)

n <- 30 * 24 * 60

dat <- arrow::read_parquet("https://github.com/mdsumner/nuyina.underway/raw/main/data-raw/nuyina_underway.parquet")


dat$longitude[dat$longitude < 0] <- -dat$longitude[dat$longitude < 0] 
print(range( dat$date_time_utc))
dat <- tibble::as_tibble(dat)
dat <- tail(dat, n)
dat$date_time_utc <- as.POSIXct(dat$date_time_utc, "%Y/%m/%d %H:%M:%S", tz = "UTC")

track <- cbind(dat$longitude, dat$latitude)

track <- terra::project(as.matrix(track), to = pcrs, from = "OGC:CRS84")
lines(track, col = "hotpink")
pt <- tail(track[!is.na(track[,1]) & !is.na(track[,2]), ], 1L)
points(pt, pch = "+", col = "hotpink")
## key locations just defined in the source of this document
pl[c("X", "Y")] <- terra::project(cbind(pl$x, pl$y), to = pcrs, from = "OGC:CRS84")
points(pl$X, pl$Y, pch = 19, col = "hotpink", cex = 0.5)
bx <- c(range(ptrack[,1], na.rm = TRUE), range(ptrack[,2], na.rm = TRUE))
rect(bx[1], bx[3], bx[2], bx[4])
claims <- terra::project(terra::vect("data-raw/claims/claim_boundaries_ps.shp"), pcrs)

plot(claims, add = TRUE)

map <- terra::vect("data-raw/CGAZ.fgb")
plot(ptrack, type = "n", asp = 1, axes = F, xlab = "", ylab = "")
title(paste0(as.Date(range(dat$date_time_utc)),collapse = ","), col.main = "white")
#plotRGB(r, add = TRUE)
ximage::ximage(r, add = TRUE)
lines(ptrack, col = "hotpink")
plot(terra::project(map, pcrs), add = TRUE, border = "#777777")

plot(claims, add = TRUE)
points(pl$X, pl$Y, pch = 19, col = "hotpink", cex = 1)
```


```{r traceplots}

vars <- c("port_solar_irradiance", "shipnav_ground_course", "air_pressure_trend3h", "fore_2_wind_from_direction_true", "port_air_temperature", "longitude", "latitude")
which(vars %in% names(dat))
 for (i in seq_along(vars)) {
   bad <- is.na(dat[[vars[i]]])
   if (any(!bad)) {
     dat2 <- dat[!bad, ]
      plot(dat2$date_time_utc, dat2[[vars[i]]], pch = 19, cex = .2, xlab = "", main = vars[i])
   }
 }

```


This is 25km sea ice concentration from NSIDC, reprojected from images published by NOAA at https://noaadata.apps.nsidc.org/NOAA/G02135/ (the projection is Transverse Mercator with central longitude 147). 

The point (and track if available) is the recent position of the [Nuyina research vessel](https://www.antarctica.gov.au/nuyina/) taken from the underway data that measures atmospheric and water properties. 

Files in 'data-raw/' contain the actual metadata and scripts. This runs as a daily task on github actions. 


Now zoom in on the ship some more.  This should just show coarse google imagery, with better Maxar imagery when near coastline.  

```{r zoom, fig.dim = c(9, 9)}
loc <- ptrack[nrow(ptrack),,  drop = FALSE]
#loc <- pl[2, c("X", "Y")]
#loc <- terra::project(cbind(-70.933004,-10.7192677)[,2:1, drop = F], to = pcrs, from = "OGC:CRS84")
xr <- loc[1,1] + c(-1000, 1000) * 34
yr <- loc[1,2] + c(-1000, 1000)  * 34


goog <- spatial.datasources::wms_googlehybrid_tms()
esri <- spatial.datasources::wms_arcgis_mapserver_ESRI.WorldImagery_tms()
gmap <- vapour::gdal_raster_image(goog, target_ext = c(xr, yr), target_crs = pcrs, target_dim = c(1024, 0))
if (length(unique(gmap[[1]])) < 800) {

xr <- loc[1,1] + c(-1000, 1000) * 800
yr <- loc[1,2] + c(-1000, 1000) * 800
gmap <- vapour::gdal_raster_image(esri, target_ext = c(xr, yr), target_crs = pcrs, target_dim = c(1024, 0))
#length(unique(gmap[[1]]))
}


plot(loc[,1:2, drop = F], xlim = xr, ylim =yr, asp = 1, axes = FALSE, xlab = "", ylab = "")
ximage::ximage(gmap, add = T)
plot(claims, add = TRUE)

lines(ptrack, col = "hotpink")
points(pl$X, pl$Y, pch = 19, col = "hotpink", cex = 0.5)


```


A sentinel-2-l2a image around the ship. 

```{r sentinel-zoom}
## there's an artefact uploaded for each run, but we should probably put these elswhere ...WIP
r <- try(rast("/vsicurl/https://github.com/mdsumner/seaice.map/raw/main/data-raw/sentinel-image.tif"))
if (!inherits(r, "try-error")) {
plotRGB(r)
lines(ptrack, col = "hotpink")
points(pl$X, pl$Y, pch = 19, col = "hotpink", cex = 0.5)
}
```



## Code of Conduct
  
Please note that the seaice.map project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
