---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
op <- options(warn = -1)

```

# seaice.map

<!-- badges: start -->
<!-- [![R-CMD-check](https://github.com/mdsumner/seaice.map/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mdsumner/seaice.map/actions/workflows/R-CMD-check.yaml)-->
<!-- badges: end -->

The goal of seaice.map is to 

- download very latest available sea ice concentration data
- generate a global map of the data with a customizable focus (i.e. global but centred on Tasmania/East Antarctica)
- also show some contextual spatio-temporal data (i.e. recent movements of a research vessel)
- explicate how to do this all with free tools. 



```{r example, fig.dim = c(7, 16.8), results="hide"}
library(terra)
r <- rast("data-raw/seaice.png")
plotRGB(r, axes = F, maxcell = prod(dim(r)[2:1]))

points(terra::project(do.call(cbind, maps::map(plot = F)[1:2]), to = terra::crs(r), from = "OGC:CRS84"), pch = ".", col = "#777777")
title(readLines("data-raw/latestdate.txt"), line = -2, col.main = "white")


#aadcgeoserver <- "WFS:https://data.aad.gov.au/geoserver/ows?service=wfs&version=2.0.0&request=GetCapabilities"
#layer <- "underway:nuyina_underway"
#info <- vapour::vapour_layer_info(aadcgeoserver, "underway:nuyina_underway")
#n <- 12 * 24 * 60
#sql <- sprintf("SELECT * FROM \"%s\" LIMIT %i OFFSET %i", layer, n, info$count - n)
#dat <- vapour::vapour_read_fields(aadcgeoserver, sql = sql)


#info <- vapour::vapour_layer_info("data-raw/nuyina_underway.parquet")
n <- 12 * 24 * 60
#sql <- sprintf("SELECT * FROM \"%s\" LIMIT %i OFFSET %i", info$layer_names[1], n, info$count - n)

#dat <- vapour::vapour_read_fields("data-raw/nuyina_underway.parquet", sql = sql)
#dat$date_time_utc <- as.POSIXct(dat$date_time_utc, "%Y/%m/%d %H:%M:%S", tz = "UTC")

dat <- arrow::read_parquet("data-raw/nuyina_underway.parquet")
dat <- tibble::as_tibble(dat)
dat <- tail(dat, n)
track <- cbind(dat$longitude, dat$latitude)

track <- terra::project(as.matrix(track), to = terra::crs(r), from = "OGC:CRS84")
lines(track, col = "hotpink")
pt <- tail(track[!is.na(track[,1]) & !is.na(track[,2]), ], 1L)
points(pt, pch = "+", col = "hotpink")
bx <- c(range(track[,1], na.rm = TRUE), range(track[,2], na.rm = TRUE))
rect(bx[1], bx[3], bx[2], bx[4])

CGAZ <- "/vsizip//vsicurl/https://github.com/wmgeolab/geoBoundaries/raw/main/releaseData/CGAZ/geoBoundariesCGAZ_ADM0.zip"
CGAZ_sql <- "SELECT shapeGroup FROM geoBoundariesCGAZ_ADM0 WHERE shapeGroup IN ('AUS','NZL','ATA')"
map <- terra::vect(CGAZ, query = CGAZ_sql)
plot(track, type = "n", asp = 1)
title(paste0(as.Date(range(dat$date_time_utc)),collapse = ","), col.main = "white")
plotRGB(r, add = TRUE)
lines(track, col = "hotpink")
#lines(terra::project(do.call(cbind, maps::map(plot = F)[1:2]), to = terra::crs(r), from = "OGC:CRS84"), lwd = 1.5, col = "#777777")
plot(terra::project(map, terra::crs(r)), add = TRUE, border = "#777777")


par(mfrow = c(5, 1))
plot(dat$date_time_utc, dat$port_solar_irradiance, pch = 19, cex = .2)
plot(dat$date_time_utc, dat$shipnav_ground_course, pch = 19, cex = .2)
plot(dat$date_time_utc, dat$air_pressure_tend3h, pch = 19, cex = .2)
plot(dat$date_time_utc, dat$fore_2_wind_from_direction_true, pch = 19, cex = .2)
plot(dat$date_time_utc, dat$port_air_temperature, pch = 19, cex = .2)

```


This is 25km sea ice concentration from NSIDC, reprojected from images published by NOAA at https://noaadata.apps.nsidc.org/NOAA/G02135/ (the projection is Transverse Mercator with central longitude 147). 

The point (and track if available) is the recent position of the [Nuyina research vessel](https://www.antarctica.gov.au/nuyina/) taken from the underway data that measures atmospheric and water properties. 

Files in 'data-raw/' contain the actual metadata and scripts. This runs as a daily task on github actions. 

## Code of Conduct
  
Please note that the seaice.map project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
